---
title: "’Machine Learning Project"
author: "Abduallah Gamal"
date: "8/25/2019"
output: html_document
---
Inroduction :
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
Data Loading:

```{r}
library(dplyr)

```

```{r}
library(ggplot2)
library(lubridate)
library(caret)
```


```{r}
library(randomForest)
```

```{r}
library(rpart)
library(rpart.plot)
library(corrplot)

data.train<- read.csv("C:/Users/Abduallah Gamal/Downloads/pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))

data.test<- read.csv("C:/Users/Abduallah Gamal/Downloads/pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
```

Data Understandaing:

```{r}
dim(data.train)
```

Data Transformation : Convert date and add new variable (Day)

```{r}
data.train$cvtd_timestamp<- as.Date(data.train$cvtd_timestamp, format = "%m/%d/%Y %H:%M")
data.train$Day<-factor(weekdays(data.train$cvtd_timestamp)) #Add day variable
```

Exploratory Data Analysis

```{r}
table(data.train$classe)
```

```{r}
prop.table(table(data.train$classe)) 
```

```{r}
prop.table(table(data.train$user_name))
```

```{r}
prop.table(table(data.train$user_name,data.train$classe),1) 
```


```{r}
prop.table(table(data.train$user_name,data.train$classe),2)
```

```{r}
prop.table(table(data.train$classe, data.train$Day),1) 
```

```{r}
qplot(x=Day, fill=classe, data = data.train)
```

Key Insights from Exploratory Data Analysis:

```{r}
### 1.Class-A activity is the most frequently used activity (28.5%) and is most frequently used by user-Jeremy
### 2.Adelmo is the most frequent user of across acitivities (20%) but he uses Class "C" activity most frequently.
### 3.Majority of the actitivies happened during Saturday's and Classes A and B are the most frequently used activites.
```

Data Cleaning:

```{r}
#### Remove columns with NA missing values
data.train <- data.train[, colSums(is.na(data.train)) == 0]
data.test <- data.test[, colSums(is.na(data.test)) == 0] 

#### Remove columns that are not relevant to accelerometer measurements.
classe<- data.train$classe
trainRemove<- grepl("^X|timestamp|window", names(data.train))
data.train<- data.train[, !trainRemove]
trainCleaned<- data.train[, sapply(data.train, is.numeric)]
trainCleaned$classe<- classe
testRemove<- grepl("^X|timestamp|window", names(data.test))
data.test<- data.test[, !testRemove]
testCleaned<- data.test[, sapply(data.test, is.numeric)]
```

Now, the cleaned data contains 19622 observations and 53 variables for both train and test datasets
Create Train and Test data sets:

```{r}
set.seed(22519)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

Data Modelling:

```{r}
#####Indetifying significant variables:
##### We will fit a predictive model using Random Forest algorithm as it gives important variables and removes multicollinearity  and outliers. We will also use 5-fold cross validation when applying the algorithm.

controlRf <- trainControl(method="cv", 5)
rfmod<- train(classe ~., data=trainData, method="rf", trControl=controlRf, importance=TRUE, ntree=100)
rfmod
```
Accuacy of the model on Validation data set:
```{r}
predictRfmod<- predict(rfmod, testData)
confusionMatrix(testData$classe, predictRfmod)
```

```{r}
accuracy <- postResample(predictRfmod, testData$classe)
accuracy
```

```{r}
Error <- 1 - as.numeric(confusionMatrix(testData$classe, predictRfmod)$overall[1])
Error
```

So, the estimated accuracy of the model is 99.32% and the estimated out-of-sample error is 0.68%.

Predicting on Test Data Set

```{r}
result <- predict(rfmod, testCleaned[, -length(names(testCleaned))])
result
```
Appendix
Correlation Matrix

```{r}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="circle")
```


Tree Visualization

```{r}
rtree<- rpart(classe ~ ., data=trainData, method="class")
prp(rtree)
```

